---
title: "Evaluation of Bivalirudin's Effect on the INR"
author: "Andrea Fetea, Brian Gulbis, Andrea C. Hall"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: hide
    fig_caption: yes
    toc: yes
    toc_float: yes
---
```{r global_options, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r library}
library(BGTools)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(pander)
library(captioner)
panderOptions("table.split.table", Inf)
fig_nums <- captioner()
```

```{r data, echo=FALSE}
get_rds("Analysis")
```

## Results

### Demographics
```{r demographics}
x <- result_table("table", analyze.demographics, group = NULL)
pander(x)
```

### Past Medical History
```{r pmh}
x <- result_table("table", analyze.diagnosis, group = NULL)
pander(x)
```

### Anticoagulation Indications
```{r indications}
x <- result_table("table", analyze.indications, group = NULL)
pander(x)
```

### Bleeding and Thrombosis
```{r events}
x <- result_table("table", analyze.bleed, group = NULL)
pander(x)
```

### Baseline Labs
```{r labs}
x <- result_table("table", analyze.labs, group = NULL)
pander(x)
```

### Percent PTT Values
```{r percent_tx}
x <- result_table("table", analyze.ptt.perc, group = NULL)
pander(x)
```

### Bivalirudin Infusion
```{r bival}
x <- result_table("table", analyze.bival, group = NULL)
pander(x)
```

### Medications During Hospitalization
```{r meds}
x <- result_table("table", analyze.meds, group = NULL)
pander(x)
```

### Blood Products
```{r prbc}
x <- result_table("table", analyze.prbc, group = NULL)
pander(x)
```

### Reversal Agents
```{r reversal}
x <- result_table("table", analyze.reversal, group = NULL)
pander(x)
```

### Primary and Secondary Endpoints
```{r endpoints}
x <- result_table("table", analyze.coags, group = NULL)
pander(x)
```

## Figures
```{r fig_caps, echo=FALSE}
fig_nums(name = "inr_tx_ptt", caption = "INR Value at Baseline and Time of First Therapeutic aPTT", display = FALSE)
fig_nums(name = "inr_bival_stop", caption = "INR Value Prior to and 4 to 8 hours After Bivalirudin Cessation", display = FALSE)
fig_nums(name = "inr_chg_bival_start", caption = "Change in INR values at bivalirudin initiation", display = FALSE)
fig_nums(name = "inr_chg_bival_stop", caption = "Change in INR values at bivalirudin cessation", display = FALSE)
fig_nums(name = "dual_bival", caption = "Change in INR and aPTT values at bivalirudin initiation and cessation", display = FALSE)
```

```{r echo=FALSE}
cap <- "inr_tx_ptt"
fig_cap <- fig_nums(cap)
```

### `r fig_nums(cap, display = "cite")`
```{r inr_tx_ptt}
results <- select(analyze.coags, pie.id, inr.before, inr.txptt) %>%
    rename(inr1 = inr.before,
           inr2 = inr.txptt) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr1),
           !is.na(inr2)) %>%
    gather(period, inr, inr1, inr2)

cols <- colorRampPalette(brewer.pal(8, "Set1"))
num.cols <- nrow(select(results, pie.id) %>% distinct)

graph <- ggplot(data = results, aes(x = period, y = inr, group = factor(pie.id))) +
    geom_line(size = 1, aes(colour = pie.id)) +
    geom_point(size = 5, pch = 21, fill = "salmon", alpha = 0.5) +
    ggtitle(fig_cap) +
    xlab(NULL) +
    ylab("INR") +
    scale_x_discrete(labels = c("Baseline", "First Therapeutic aPTT"), expand = c(0.25, 0.25)) +
    scale_color_manual(values = cols(num.cols)) +
    # theme_bw() +
    theme(legend.position = "none")
    # theme(legend.position = "none", plot.title = element_text(size = 20), 
    #       axis.title = element_text(size = 18), axis.text = element_text(size = 16))

print(graph)
```

```{r echo=FALSE}
cap <- "inr_bival_stop"
fig_cap <- fig_nums(cap)
```

### `r fig_nums(cap, display = "cite")`
```{r inr_bival_stop}
results <- select(analyze.coags, pie.id, inr.before.stop, inr.after.stop) %>%
    rename(inr1 = inr.before.stop,
           inr2 = inr.after.stop) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr1),
           !is.na(inr2)) %>%
    gather(period, inr, inr1, inr2)

graph <- ggplot(data = results, aes(x = period, y = inr, group=factor(pie.id))) +
    geom_line(size = 1, aes(colour = pie.id)) +
    geom_point(size = 5, pch = 21, fill = "salmon", alpha = 0.5) +
    ggtitle(fig_cap) +
    xlab(NULL) +
    ylab("INR") +
    scale_x_discrete(labels = c("Prior to Cessation", "After Cessation"), expand = c(0.25, 0.25)) +
    # theme_bw() +
    theme(legend.position = "none")
    # theme(legend.position = "none", plot.title = element_text(size = 20), 
    #       axis.title = element_text(size = 18), axis.text = element_text(size = 16))

print(graph)
```

```{r echo=FALSE}
cap <- "inr_chg_bival_start"
fig_cap <- fig_nums(cap)
ylim.inr <- c(1, 3.5)
x.breaks <- c(-48, -24, 0, 24, 48, 72)
graph.alpha <- 0.3
```

### `r fig_nums(cap, display = "cite")`
```{r inr_chg_bival_start}
graph <- analyze.labs.serial %>%
    filter(lab == "inr",
           first(lab.datetime) < bival.start,
           last(lab.datetime) > bival.start,
           lab.datetime <= bival.start + days(3),
           warf.start >= bival.start + days(1)) %>%
    mutate(bival.diff = as.numeric(difftime(lab.datetime, bival.start, units = "hours"))) %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    geom_line(alpha = graph.alpha) +
    geom_smooth(aes(group = NULL), color = "blue", se = FALSE) +
    geom_smooth(aes(group = bival.diff <= 0), method = "lm", color = "black") +
    ggtitle(fig_cap) +
    xlab("Time from bivalirudin initiation (hours)") +
    ylab("INR") +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.inr) +
    theme(legend.position = "none")

print(graph)
```

Figure `r fig_nums(cap, display = "num")` is depicting the mean INR value before and after bivalirudin initation (represented by the time point 0 on the x-axis). The blue line represents the smoothed mean INR value throughout the time period. The black line represents the linear mean INR value, split for values before initiation and after initiation. The colored lines and points represent individual patient INR values. Data from patients who started warfarin during the first 24 hours after bivalirudin initiation were excluded, so the INR increase during the first 24 hours of bivalirudin infusion should be the result of bivalirudin.

```{r echo=FALSE}
cap <- "inr_chg_bival_stop"
fig_cap <- fig_nums(cap)
```

### `r fig_nums(cap, display = "cite")`
```{r inr_chg_bival_stop}
graph <- analyze.labs.serial %>%
    filter(lab == "inr",
           last(lab.datetime) > bival.stop,
           lab.datetime >= bival.stop - hours(48),
           lab.datetime <= bival.stop + days(3),
           warf.start >= bival.start + days(1)) %>%
    mutate(bival.diff = as.numeric(difftime(lab.datetime, bival.stop, units = "hours"))) %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    geom_line(alpha = graph.alpha) +
    geom_smooth(aes(group = NULL), color = "blue", se = FALSE) +
    geom_smooth(aes(group = bival.diff <= 0), method = "lm", color = "black") +
    ggtitle(fig_cap) +
    xlab("Time from bivalirudin cessation (hours)") +
    ylab("INR") +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.inr) +
    theme(legend.position = "none")

print(graph)
```

Figure `r fig_nums(cap, display = "num")` is depicting the mean INR value before and after bivalirudin cessation (represented by the time point 0 on the x-axis). The blue line represents the smoothed mean INR value throughout the time period. The black line represents the linear mean INR value, split for values before cessation and after cessation. The colored lines and points represent individual patient INR values. Data from patients who started warfarin during the first 24 hours after bivalirudin initiation were excluded.

```{r echo=FALSE}
cap <- "dual_bival"
fig_cap <- fig_nums(cap)
ylim.ptt <- c(45, 85)
```

### `r fig_nums(cap, display = "cite")`
```{r dual_bival}
graph.data <- analyze.labs.serial %>%
    filter(lab == "inr" | lab == "ptt",
           first(lab.datetime) < bival.start,
           last(lab.datetime) > bival.start,
           lab.datetime <= bival.start + days(3),
           warf.start >= bival.start + days(1)) %>%
    mutate(bival.diff = as.numeric(difftime(lab.datetime, bival.start, units = "hours"))) 

graph.inr1 <- filter(graph.data, lab == "inr") %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    # geom_line(alpha = graph.alpha) +
    # geom_smooth(aes(group = bival.diff <= 0), method = "lm", color = "black") +
    geom_smooth(aes(group = NULL), color = "black") +
    ggtitle("A. INR and bivalirudin initiation") +
    xlab(NULL) +
    ylab("INR") +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.inr) +
    theme(legend.position = "none")

graph.ptt1 <- filter(graph.data, lab == "ptt") %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    # geom_line(alpha = graph.alpha) +
    geom_smooth(aes(group = NULL), color = "black") +
    ggtitle("C. aPTT and bivalirudin initiation") +
    xlab("Time from bivalirudin initiation (hours)") +
    ylab("aPTT (seconds)") +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.ptt) +
    theme(legend.position = "none")

graph.data <- analyze.labs.serial %>%
    filter(lab == "inr" | lab == "ptt",
           last(lab.datetime) > bival.stop,
           lab.datetime >= bival.stop - hours(48),
           lab.datetime <= bival.stop + days(3),
           warf.start >= bival.start + days(1)) %>%
    mutate(bival.diff = as.numeric(difftime(lab.datetime, bival.stop, units = "hours")))

graph.inr2 <- filter(graph.data, lab == "inr") %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    # geom_line(alpha = graph.alpha) +
    geom_smooth(aes(group = NULL), color = "black") +
    ggtitle("B. INR and bivalirudin cessation") +
    xlab(NULL) +
    ylab(NULL) +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.inr) +
    theme(legend.position = "none")

graph.ptt2 <- filter(graph.data, lab == "ptt") %>%
    ggplot(aes(x = bival.diff, y = result, color = pie.id)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(alpha = graph.alpha) +
    # geom_line(alpha = graph.alpha) +
    geom_smooth(aes(group = NULL), color = "black") +
    ggtitle("D. aPTT and bivalirudin cessation") +
    xlab("Time from bivalirudin cessation (hours)") +
    ylab(NULL) +
    scale_x_continuous(breaks = x.breaks) +
    coord_cartesian(ylim = ylim.ptt) +
    theme(legend.position = "none")

grid.arrange(graph.inr1, graph.inr2, graph.ptt1, graph.ptt2, nrow = 2, ncol = 2, top = fig_cap)
```

