---
title: "Exploratory Analysis"
author: "Andrea Fetea, Brian Gulbis, Christie Hall"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
source("../src/05_make-analysis-datasets.R")
library(pander)
panderOptions('table.split.table', Inf)
```

## Patients

* Patients receiving bivalirudin: `r nrow(pts.identified)`
* Patients receiving warfarin: `r nrow(tmp.warf.start)`
* Patients with at least one overlapping dose: `r nrow(tmp.overlap)`
* Patients meeting inclusion criteria: `r nrow(pts.eligible)`
* Patients excluded:
    - Alternate anticoagulant during transition interval: `r nrow(excl.alt.anticoag)`
    - Bleeding within 48 hours prior to bivalirudin: `r nrow(excl.bleed)`
    - Diagnosis of liver disease or hypercoaguable condidtion: `r nrow(excl.diag)`
    - INR >1.5 at time of bivalirudin initiation: `r nrow(excl.inr)`
    - Vitamin K with 5 days of bivalirudin: `r nrow(excl.vitk)`
    - Thrombolytics within 48 hours of bivalirudin: `r nrow(excl.tpa)`
    - Major inducers: `r nrow(excl.inducer)`
    - Pregnant: `r nrow(excl.preg)`
* Patients included in study: `r nrow(pts.include)`

### Exclusion Criteria Variables

* Hours of bivalirudin and warfarin overlap: `r ec.overlap`
* Days prior to bivalirudin with an INR < 1.5: `r ec.inr`

## Demographics

```{r echo=FALSE}
results <- select(data.demograph, age:disposition, bival.duration:weight, proc.48hrs) %>%
    mutate(proc.48hrs = as.factor(proc.48hrs))
cont <- sapply(results, is.numeric)
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)

## check for normal distribution for continuous variables
```

## Diagnosis

```{r echo=FALSE}
results <- ungroup(data.pmh) %>% 
    select(-pie.id) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

### Anticoagulation Indications

As specified in the warfarin goals

```{r echo=FALSE}
results <- ungroup(data.warf) %>% 
    select(-pie.id, -indication) %>%
    mutate_each(funs(as.factor)) 
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

### Bleeding

```{r echo=FALSE}
results <- ungroup(data.bleed.bival) %>%
    select(major.bleed, minor.bleed, hgb.drop, bival.drop.diff, warf.drop.diff) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## New Thrombosis
```{r echo=FALSE}
results <- ungroup(data.new.thrmb) %>% 
    select(manual) %>%
    rename(new.thromb = manual) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Baseline Labs

Most recent lab value prior to bivalirudin initiation (except for the heparin antibody ELISA, which could have been performed at any time)

```{r echo=FALSE}
results <- select(data.labs.baseline, -pie.id) %>%
    mutate(elisa = as.factor(elisa))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Serial Lab Values

Summary of labs from 48 hours prior to bivalirudin initiation to 5 days after bival cesation

```{r echo=FALSE}
results <- ungroup(data.labs.serial) %>%
    select(-pie.id) %>%
    group_by(lab) %>%
    do(result = lapply(as.list(.[,-1]), Summarize))
names(results$result) <- results$lab
pander(results$result)
```

## Bivalirudin Infusion

```{r echo=FALSE}
results <- select(data.bival, -pie.id)
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Medications During Hospitalization

```{r echo=FALSE}
results <- select(data.meds, -pie.id) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Blood Products

```{r echo=FALSE}
results <- select(data.blood, -pie.id) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Reversal Agents

```{r echo=FALSE}
results <- select(data.reversal, -pie.id) %>%
    mutate_each(funs(as.factor))
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

## Primary and Secondary Endpoints

```{r echo=FALSE}
results <- select(data.labs.coags, -pie.id)
ds1 <- lapply(as.list(results), Summarize)
pander(ds1)
```

```{r echo=FALSE}
graph <- ggplot(data=results, aes(x=inr.change)) +
    geom_density()
graph
```

```{r echo=FALSE}
graph <- ggplot(data=results, aes(x=inr.change.stop)) +
    geom_density()
graph
```

```{r echo=FALSE, results="hide"}
# fig.nums <- captioner()
# fig.nums("inr.start", )
# fig.nums("inr.stop", )
# fig.nums("inr.change", )
```

```{r echo=FALSE, fig.cap="Figure 1. INR at bivalirudin initiation and first therapeutic aPTT in paired samples"}
results <- select(data.labs.coags, pie.id, inr.before, inr.txptt) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr.before),
           !is.na(inr.txptt)) %>%
    gather(period, inr, inr.before, inr.txptt)

graph <- ggplot(data=results, aes(x=period, y=inr, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme_bw() +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="Figure 2. INR just prior to bivalirudin cesation and between 4 and 8 hours after cessation in paired samples"}
results <- select(data.labs.coags, pie.id, inr.before.stop, inr.after.stop) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr.before.stop),
           !is.na(inr.after.stop)) %>%
    gather(period, inr, inr.before.stop, inr.after.stop)

graph <- ggplot(data=results, aes(x=period, y=inr, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme_bw() +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="Figure 3. Absolute INR change at bivalirudin initiation and cesation"}
results <- select(data.labs.coags, inr.change, inr.change.stop) %>%
    rename(bival.start = inr.change,
           bival.stop = inr.change.stop) %>%
    mutate(bival.stop = abs(bival.stop)) %>%
    gather(period, change) %>%
    filter(!is.na(change))

graph <- ggplot(data=results, aes(x=period, y=change)) +
    geom_boxplot() + 
    theme_bw()
    # theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="INR change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, inr.change, inr.change.stop) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr.change),
           !is.na(inr.change.stop)) %>%
    gather(period, change, inr.change, inr.change.stop)

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="INR percent change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, inr.change.prct, inr.change.stop.prct) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr.change.prct),
           !is.na(inr.change.stop.prct)) %>%
    gather(period, change, inr.change.prct, inr.change.stop.prct)

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="Comparison of the amount of INR change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, inr.change.prct, inr.change.stop.prct) %>%
    group_by(pie.id) %>%
    filter(!is.na(inr.change.prct),
           !is.na(inr.change.stop.prct)) %>%
    gather(period, change, inr.change.prct, inr.change.stop.prct) %>%
    mutate(change = abs(1 - change))

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="PTT change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, ptt.change, ptt.change.stop) %>%
    group_by(pie.id) %>%
    filter(!is.na(ptt.change),
           !is.na(ptt.change.stop)) %>%
    gather(period, change, ptt.change, ptt.change.stop)

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="PTT percent change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, ptt.change.prct, ptt.change.stop.prct) %>%
    group_by(pie.id) %>%
    filter(!is.na(ptt.change.prct),
           !is.na(ptt.change.stop.prct)) %>%
    gather(period, change, ptt.change.prct, ptt.change.stop.prct)

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

```{r echo=FALSE, fig.cap="Comparison of the amount of PTT change at bivalirudin initiation and cesation in paired samples"}
results <- select(data.labs.coags, pie.id, ptt.change.prct, ptt.change.stop.prct) %>%
    group_by(pie.id) %>%
    filter(!is.na(ptt.change.prct),
           !is.na(ptt.change.stop.prct)) %>%
    gather(period, change, ptt.change.prct, ptt.change.stop.prct) %>%
    mutate(change = abs(1 - change))

graph <- ggplot(data=results, aes(x=period, y=change, group=factor(pie.id))) +
    geom_line(size=1, aes(colour=pie.id)) +
    geom_point(size=5, pch=21, fill="salmon", alpha=0.5) +
    theme(legend.position="none")
graph
```

Correlation between INR changes and PTT changes

```{r echo=FALSE}
results <- ungroup(data.labs.coags) %>% select(inr.change.prct, inr.change.stop.prct, ptt.change.prct, ptt.change.stop.prct, -pie.id) 
ds1 <- cor(results, use="pairwise.complete.obs")
pander(ds1)
```
